---
title: "R Notebook for image classification"
output: html_notebook
---

Those are my notes to understand how to buld an image classifier in R. I am using the dogs and cats images from the Kaggle competition: https://www.kaggle.com/c/dogs-vs-cats and following this tutorial by  Dr. Bharatendra Rai: https://www.youtube.com/watch?v=iExh0qj2Ouo

First, loading the libraries:
```{r}
library(tensorflow) #machine learning
library(keras) #neural network
library(EBImage) #image processing
```

Setting up the environment and reading images: 
```{r}
setwd("/home/caroline/projects/dogs-vs-cats/data/raw/train")

pics <- c("cat.1.jpg", "cat.2.jpg", "cat.3.jpg", "cat.4.jpg", "cat.5.jpg", "cat.6.jpg", "dog.1.jpg", "dog.2.jpg", "dog.3.jpg", "dog.4.jpg", "dog.5.jpg", "dog.6.jpg")

mypic <- list()

for(i in 1:12) {
  mypic[[i]]<-readImage(pics[i])
}
```

Explore and dispplay images
```{r}
print(mypic[[1]])
display(mypic[[1]])
summary(mypic[[1]])
hist(mypic[[1]])
str(mypic)
```

Resize all pictures to have all dimensions
```{r}
for (i in 1:12) {mypic[[i]] <- resize(mypic[[i]], 28, 28)}
```

Explore resized images
```{r}
class(mypic)
print(mypic[[1]])
display(mypic[[1]])
summary(mypic[[1]])
hist(mypic[[1]])
str(mypic)
```

Reshape all pictures to a single array
```{r}
for (i in 1:12) {mypic[[i]] <- array_reshape(mypic[[i]], c(28, 28, 3))}
str(mypic)
class(mypic)
```

Row Bind
```{r}
trainx <- NULL
for (i in 7:11) {trainx <- rbind(trainx, mypic[[i]])}
str(trainx)
testx <- rbind(mypic[[6]], mypic[[12]])
trainy <- c(0,0,0,0,0,1,1,1,1,1 )
testy <- c(0, 1)
```

One Hot Encoding
```{r}
trainLabels <- to_categorical(trainy)
testLabels <- to_categorical(testy)
```

Create the Model
```{r}
model <- keras_model_sequential()
model %>%
         layer_dense(units = 256, activation = 'relu', input_shape = c(2352)) %>%
         layer_dense(units = 128, activation = 'relu') %>%
         layer_dense(units = 2, activation = 'softmax')
summary(model)
```

Compile the model
```{r}
model %>%
         compile(loss = 'binary_crossentropy',
                 optimizer = optimizer_rmsprop(),
                 metrics = c('accuracy'))
```

Fit the Model
```{r}
history <- model %>%
         fit(trainx,
             trainLabels,
             epochs = 30,
             batch_size = 32,
             validation_split = 0.2)
```

Evaluation & Prediction - train data
```{r}
model %>% evaluate(---, ---)
pred <- model %>% predict_classes(trainx)
table(Predicted = pred, Actual = trainy)
prob <- model %>% predict_proba(trainx)
cbind(prob, Prected = pred, Actual= trainy)
```{r}